Homework 8 - Linear Regression
#Q1 - Plot
```{r}
#get variables in
age <- c(3,4,5,6,7,8,9,11,12,14,15,16,17)
wing_length <- c(1.4,1.5,2.2,2.4,3.1,3.2,3.2,3.9,4.1,4.7,4.5,5.2,5)

#standard plot
plot(age,wing_length)


```
#Q2 - calculate and plot regression line
```{r}
#using built in R function to calculate regression
regression <- lm(wing_length ~ age)

#then plot it
plot(age,wing_length, main = 'Wing Length vs Age', xlab = "Age", ylab = "Wing Length", pch = 19, col = "blue")
#and add the regression line
abline(regression, col = 'red')
```
#Q3 - reject Ho?
Yes because as you will see from the F-statistic row at the bottom, the p-value is 3.01e-9 which is far less than 0.05.
Therefore there is significance and we can reject that the slope is 0.
```{r}
#get a list of the regression stats
calculated_stats <- summary(regression)

#and print them out to see the p-value at the bottom based on the F statistic
print(calculated_stats)


```
#Q4 - CI
See the R console tab for the 95% CI around the slope.
```{r}
#use built in R CI function specifically built around the linear model function used in question 2
confidence_interval <- confint(regression, "age", level = 0.95)

#print
print(paste("95% CI is","[",confidence_interval[1],",",confidence_interval[2],"]"))


#to visualize, lets put the data in a dataframe
df <- data.frame("age" = age, "wing_length" = wing_length)

#then using ggplot, plot the data
library(ggplot2)
ggplot(df, aes(x = age, y = wing_length)) +
  geom_point() +  # Add scatterplot points
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # Add regression line with CI
  labs(
    title = "Scatterplot with Regression Line and 95% CI",
    x = "Age",
    y = "Wing Length"
  )


```
#Q5 - r squared
```{r}
#r squared was already calculated from the R regression function so just extract it
print(paste("The r^2 is",calculated_stats$r.squared))
```
#Q6 - pearson's R
```{r}
#from a previous homework calculate pearson's r using the built in R function
pearsons_cor <- cor(x = age, y = wing_length)

print(paste("Pearson's r is", pearsons_cor))
```
#Q7 - add some noise to the data
At the suggestion of ChatGPT, we can add noise to each point by using the rnorm function to generate random numbers from a normal distribution.
With a high noise rate (2.3 standard deviations), the trend is essentially gone with a non-significant p value (most of the time bc this depends on randomly generated noise).
95% CI is now a much wider range.
```{r}
#set an arbitrary noise level which will be the standard deviation
noise_level <- 2.3

#calculate new noisy ages
age2 <- age + rnorm(length(age), mean = 0, sd = noise_level)
wing_length2 <- wing_length + rnorm(length(wing_length), mean = 0, sd = noise_level)

#get everything into a dataframe
noisy_data <- data.frame("age" = age2, "wing_length" = wing_length2)

#calculate regression statistics
regression2 <- lm(wing_length ~ age, data = noisy_data)

#print the stats to see the new p value
print(summary(regression2))

#then plot with CI as before
library(ggplot2)
ggplot(noisy_data, aes(x = age, y = wing_length)) +
  geom_point() +  # Add scatterplot points
  geom_smooth(method = "lm", se = TRUE, color = "blue") +  # Add regression line with CI
  labs(
    title = "Noisy Scatterplot with Regression Line and 95% CI",
    x = "Age",
    y = "Wing Length"
  )

#and recalculate r and r^2
print(paste("Pearson's r is now",cor(x = noisy_data$age, y = noisy_data$wing_length)))
noisy_cal_stats <- summary(regression2)
print(paste("R squared is now",noisy_cal_stats$r.squared))

```

