Homework 6: Multiple Comparisons
#Simulate p values
```{r}

#create a list of 1000 numbers to represent how many tests we want to run
experiment_iterations <- seq(from = 1, to = 1000, by = 1)

#create an empty list for the p values
final_pvalues <- c()
#specify alpha
alpha <- 0.05

#loop through 1000 replications
for (i in experiment_iterations) {
  
#generate data sample 1
data1 <- rnorm(30, mean = 4, sd = 3)
#generate data sample 2
data2 <- rnorm(30, mean = 4, sd = 3)

#run a non-paired 2 sample t test
  placeholder <- t.test(x = data1, y = data2, paired = FALSE, var.equal = TRUE)
  #extract just the p value
  statistic <- placeholder$p.value
  
  #add that p value to the list created at the beginning
  final_pvalues <- append(statistic, final_pvalues)
  
}

#make a dataframe with all the p values
total_pvalues <- data.frame("experiment_iteration" = experiment_iterations, "p_values" = final_pvalues)
#count all the p values that are significant
count <- as.numeric(sum(total_pvalues$p_values < alpha))

#and then print that summation
print(paste(count,"trials identified as <",alpha))

```
#Multiple comparison correction
```{r}

#Bonferroni method

#calculate adjusted alpha
adjusted_alpha <- alpha/1000

#using the same dataframe of simulated p values, count how many are significant now, based on the adjusted alpha
bonfer_count <- as.numeric(sum(total_pvalues$p_values < adjusted_alpha))

#and then print that summation
print(paste(bonfer_count,"trials identified as significant with the Bonferroni method"))



#Benjamini-Hochberg

#load the simple FDR package, created, written, and published to CRAN by yours truly
#I built this package specifically around the Benjamini-Hochberg procedure
library(simpleFDR)

#calculate significance based on the BH method using the simFDR function and return all values, not filtered by significance
BH_adjustments <- simFDR(total_pvalues, sig_only = FALSE)

#count how many are significant now, which is done by counting how many NAs there are not (significance is denoted by an asterisk)
BH_count <- as.numeric(sum(!is.na(BH_adjustments$significant)))

#and print that number
print(paste(BH_count,"trials identified as significant with the Benjamini-Hochberg method"))


```
#Rerun the full script with means 1 & 2, and then more means with larger differences
As you will see from the resulting plot, when there is a difference of 1 between the means, the number of significanes out of 1000 experiments increases (compared to equal means) which drops substantially with the bonferroni correction, but a less significant drop with the BH correction.  For larger differences between the means, this same pattern holds, where the BH correction does not substantially effect the original number of significances, but the Bonferroni does, but eventually with large enough differences, all 1000 trials are significant even with both correction methods.
```{r}
#create a list of 1000 numbers to represent how many tests we want to run
experiment_iterations <- seq(from = 1, to = 1000, by = 1)

#create different mean sequences
mean1seq <- seq(from = 1, to = 10, by = 1)
mean2seq <- seq(from = 2, to = 20, by = 2)

#create a sequence to cycle through 10 different mean difference scenarios
mean_difference_iterations <- seq(from = 1, to = 10, by = 1)

#create empty vectors for counts of significances under all 3 methods
sig_no_correction <- c()
sig_bonferroni <- c()
sig_BH <- c()

#repeat the script so far looping though it for each of the 10 mean differences scenarios
for (j in mean_difference_iterations) {
  


#create an empty list for the p values
final_pvalues <- c()
#specify alpha
alpha <- 0.05

#loop through 1000 replications
for (i in experiment_iterations) {
  
#generate data sample 1
data1 <- rnorm(30, mean = mean1seq[j], sd = 3)
#generate data sample 2
data2 <- rnorm(30, mean = mean2seq[j], sd = 3)

#run a non-paired 2 sample t test
  placeholder <- t.test(x = data1, y = data2, paired = FALSE, var.equal = TRUE)
  #extract just the p value
  statistic <- placeholder$p.value
  
  #add that p value to the list created at the beginning
  final_pvalues <- append(statistic, final_pvalues)
  
}

#make a dataframe with all the p values
total_pvalues <- data.frame("experiment_iteration" = experiment_iterations, "p_values" = final_pvalues)
#count all the p values that are significant
count <- as.numeric(sum(total_pvalues$p_values < alpha))






#corrections

#Bonferroni method

#calculate adjusted alpha
adjusted_alpha <- alpha/1000

#using the same dataframe of simulated p values, count how many are significant now, based on the adjusted alpha
bonfer_count <- as.numeric(sum(total_pvalues$p_values < adjusted_alpha))




#Benjamini-Hochberg

#load the simple FDR package, created, written, and published to CRAN by yours truly
#I built this package specifically around the Benjamini-Hochberg procedure
library(simpleFDR)

#calculate significance based on the BH method using the simFDR function and return all values, not filtered by significance
BH_adjustments <- simFDR(total_pvalues, sig_only = FALSE)

#count how many are significant now, which is done by counting how many NAs there are not (significance is denoted by an asterisk)
BH_count <- as.numeric(sum(!is.na(BH_adjustments$significant)))



#add the counts for each method to the appropriate vector created at the beginning
sig_no_correction <- append(count, sig_no_correction)
sig_bonferroni <- append(bonfer_count, sig_bonferroni)
sig_BH <- append(BH_count, sig_BH)

}

#reverse those counts to get them in the correct order
sig_no_correction <- rev(sig_no_correction)
sig_bonferroni <- rev(sig_bonferroni)
sig_BH <- rev(sig_BH)

#add all the counts as a function of the difference between the 2 means
mean_difference_dataframe <- data.frame("mean_difference" = mean_difference_iterations, "original_significances" = sig_no_correction, "bonferroni_significances" = sig_bonferroni, "BH_significances" = sig_BH)

#plot the number of significances out of 1000 for each method
plot(mean_difference_dataframe$mean_difference, mean_difference_dataframe$original_significances, main = "Difference between means vs Number of Significances out of 1000 tries", xlab = "Difference between the 2 means", ylab = "number of signifances", xaxt = "n", col = "blue", type = "l", ylim = c(1,1000))
axis(1, at = 1:10)

lines(mean_difference_dataframe$mean_difference, mean_difference_dataframe$bonferroni_significances, col = "red")
lines(mean_difference_dataframe$mean_difference, mean_difference_dataframe$BH_significances, col = "green")

points(mean_difference_dataframe$mean_difference, mean_difference_dataframe$original_significances, col = "blue")
points(mean_difference_dataframe$mean_difference, mean_difference_dataframe$bonferroni_significances, col = "red")
points(mean_difference_dataframe$mean_difference, mean_difference_dataframe$BH_significances, col = "green")

legend("bottomright", legend = c("original significances", "bonferroni corr", "BH corr"), 
       col = c("blue", "red", "green"), lty = 1, pch = 19)



```

